% !TeX root = ../thesis.tex
\chapter{Exercise Tools as a Solution Direction}
The previous chapter established that the tools currently used for arithmetic research do their job as research instruments but fail as experiences for the children who use them. They lack the motivational, adaptive, and visual qualities that keep children engaged across multiple sessions, and the programming barrier makes it impractical for researchers to add these qualities themselves. The question that follows naturally is: where do these qualities already exist?\\ \\
The answer is not far away. Across primary schools in Belgium and the Netherlands, children already use digital tools for arithmetic practice on a daily basis. Platforms like Automatus and Rekentuin are designed from the ground up to be engaging, adaptive, and easy for children to use. They look like games, they feel like games, and children are willing to spend time on them without being asked twice. In other words, these tools already solve many of the problems identified in Section 2.4: they offer the ecological validity, the sustained motivation, and the familiarity that research tools lack.\\ \\
This chapter examines exercise tools as a possible solution direction. It begins by defining what makes a good exercise tool, then walks through two prominent examples in the field, and ends by asking the inevitable follow-up question: if exercise tools solve the engagement problem, can they also serve as research tools? The answer, as will become clear, is not quite.

\section{What a good exercise tool needs}
Where Section 2.2 asked what a research tool needs, this section asks the same question for the other side: what makes a good tool for arithmetic practice in the classroom? The properties listed below were developed through discussions with primary school teachers, educational researchers, and the co-supervisor of this thesis, and reflect what practitioners in the field consistently identify as important. Where possible, connections to the educational psychology literature are drawn to ground these observations in theory.

\subsection*{Adaptivity}
Perhaps the single most important property of a good exercise tool is that it adapts to the child using it. Not every child in a classroom is at the same level, and a tool that presents the same exercises to everyone will inevitably bore the strongest students and frustrate the weakest ones. This idea has deep theoretical roots. Vygotsky's concept of the zone of proximal development describes the range of tasks that a child cannot yet do independently but can accomplish with appropriate support \cite{vygotsky1978mind}. A good adaptive system operationalises this concept by selecting exercises that sit within the child's zone: challenging enough to promote learning, but not so hard that the child gives up. Csikszentmihalyi's flow theory makes a closely related point from the perspective of engagement: when the difficulty of a task is well-matched to the learner's skill level, the learner enters a state of deep concentration and intrinsic motivation; when the balance is off, the result is either anxiety or boredom \cite{csikszentmihalyi1990flow}. An exercise tool that does not adapt is, in effect, guaranteeing that most of the children using it will be outside their flow channel most of the time.

\subsection*{Engagement}
Adaptivity keeps the difficulty right, but it does not on its own make a child want to open the tool again tomorrow. Engagement is about everything else that makes the experience feel worthwhile: a story that gives context to the exercises, visual rewards for progress, puzzles to unlock, a character to interact with. These are gamification elements, and while the term is sometimes used loosely, systematic reviews of gamification in education have consistently found that elements such as points, levels, narrative framing, and immediate feedback can improve motivation and engagement in mathematics \cite{manzanoleon2021gamification}. For an exercise tool aimed at primary school children, engagement is not optional. A tool that children do not want to use is a tool that does not get used, regardless of how well-designed its exercises are.

\subsection*{Feedback}
Children need to know whether they got the answer right, and ideally they need to know it immediately. Delayed or absent feedback leaves the child uncertain about whether they are learning, and it removes one of the most natural drivers of improvement: the desire to correct a mistake you just made. Good feedback goes beyond a simple correct/incorrect signal. It can show the right answer when the child was wrong, acknowledge a good self-assessment when the child judged their own confidence accurately, or offer encouragement after a difficult exercise. The key is that the feedback is timely, clear, and meaningful.

\subsection*{Personalisation}
Personalisation is related to adaptivity but distinct from it. Where adaptivity adjusts the difficulty automatically based on performance, personalisation gives the child the feeling that the tool is theirs. This has two sides. The first is functional: letting children choose which type of exercise to practise, whether they want to see a timer, or what difficulty level they prefer. This kind of choice can foster a sense of ownership and metacognitive awareness. For example, when a child decides to practise multiplication because they know it is their weakest area, they are doing something more valuable than just answering another question. The second side is personal identity: an avatar, a nickname, a progress tracker that shows their journey, a customised home screen. These elements may seem cosmetic, but they transform a generic program into something that feels like it belongs to the child, which in turn makes the child more likely to come back to it.

\subsection*{Accessibility}
An exercise tool is only useful if children can actually reach it. In practice this means the tool should run in a web browser, work on the devices that schools actually have (laptops, tablets, Chromebooks), and not require any software installation. It also means the tool should be available outside the classroom, so that children can practise at home if they want to. A tool that only works on a specific operating system, or that requires a local installation managed by an IT department, will inevitably reach fewer children than one that runs on any device with a browser.

\subsection*{User friendliness}
Finally, the tool has to be usable by its target audience, which in this case is children between roughly seven and eleven years old. This means large buttons, clear icons, simple navigation, and a visual language that children understand without needing instructions. Cognitive load theory reminds us that every unnecessary interface element competes for the limited working memory resources that the child should be spending on the actual exercise \cite{sweller1988cognitive}. A well-designed exercise tool is one where the interface disappears: the child thinks about the arithmetic, not about how to operate the tool.

\section{Existing execise tools: Automatus and Rekentuin}

\section{Why exercise tools alone are not the answer}
You maybe already ask? but what with the research properties?
\section{The gap: no existing tool covers both sides}